---
title: "Summary stats"
output:
  word_document: default
  html_document: default
---

```{r}
# Load libraries
library(tidyverse)
library(knitr)
library(here)
```

```{r}
# Read in all the specimen data
specimens <- read_csv(here("data/all-specimen-data-2021-07.csv")) 
specimens_all <- read_csv(here("data/all-specimen-data-unsexed-2021-07.csv")) 
```

## How many male, female and unsexed specimens are there?

```{r}
  sum1 <-
  specimens_all %>%
  add_count(class, name = "n") %>%
  add_count(class, sex, name = "nn") %>%
  select(class, sex, n, nn) %>%
  distinct() %>%
  mutate('%' = round(nn/n*100, 2))

  kable(sum1)
```

## Considering only sexed specimens, how many male and female specimens are there?

```{r}
  sum2 <-
  specimens %>%
  filter(!is.na(sex)) %>%
  add_count(class, name = "n") %>%
  add_count(class, sex, name = "nn") %>%
  select(class, sex, n, nn) %>%
  distinct() %>%
  mutate('%' = round(nn/n*100, 2))

  kable(sum2)
```

## What is the median proportion of females per species?

```{r}
  proportion <-
  specimens %>%
  add_count(class, binomial, name = "n") %>%
  add_count(class, binomial, sex, name = "nn") %>%
  select(class, binomial, sex, n, nn) %>%
  distinct() %>%
  mutate(percent = round(nn/n*100, 2)) %>%
  complete(sex, nesting(class, binomial),
           fill = list(nn = 0, percent = 0)) %>%
  filter(sex == "Female") %>%
  group_by(class) %>%
  summarise('median % female' = median(percent))

  kable(proportion)
```

## What is the median proportion of females per species for those with >= 10 specimens only?

```{r}
  proportion2 <-
  specimens %>%
  select(class, binomial, sex) %>%
  add_count(class, binomial, name = "n") %>%
  add_count(class, binomial, sex, name = "nn") %>%
  # Add in data for species with no males or no females
  complete(sex, nesting(class, binomial, n), fill = list(nn = 0)) %>%
  distinct() %>%
  mutate(percent = round(nn/n*100, 2)) %>%
  filter(sex == "Female" & n >= 10) %>%
  group_by(class) %>%
  summarise('median % female' = median(percent))

  kable(proportion2)
```

## And what if we only look at species with > 10 specimens

```{r}
  sum3 <-
  specimens %>%
  add_count(binomial, name = "spec") %>%
  filter(spec >= 10) %>%
  add_count(class, name = "n") %>%
  add_count(class, sex, name = "nn") %>%
  select(class, sex, n, nn) %>%
  distinct() %>%
  mutate('%' = round(nn/n*100, 2))

  kable(sum3)
```

## How many species are there?
```{r}
  sum4 <-
  specimens %>%
  add_count(binomial, name = "spec") %>%
  filter(spec >= 10) %>%
  select(class, binomial) %>%
  distinct() %>%
  group_by(class) %>%
  summarise(n())

  kable(sum4)
```

## Which are the worst contenders?
### Amphibians
```{r}
  worst <-
  specimens %>%
  add_count(class, binomial, name = "n") %>%
  add_count(class, binomial, sex, name = "nn") %>%
  select(class, binomial, sex, n, nn) %>%
  # Add in data for species with no males or no females
  complete(sex, nesting(class, binomial, n), fill = list(nn = 0)) %>%
  distinct() %>%
  mutate(percent = round(nn/n*100, 2)) %>%
  filter(sex == "Female" & n >= 10) %>%
  arrange(class, percent)

kable(slice(worst, 1:10))
```

```{r}
  best <-
  specimens %>%
  add_count(class, binomial, name = "n") %>%
  add_count(class, binomial, sex, name = "nn") %>%
  select(class, binomial, sex, n, nn) %>%
  # Add in data for species with no males or no females
  complete(sex, nesting(class, binomial, n), fill = list(nn = 0)) %>%
  distinct() %>%
  mutate(percent = round(nn/n*100, 2)) %>%
  filter(sex == "Female" & n >= 10) %>%
  arrange(class, -percent)

kable(slice(best, 1:10))
```

### Reptiles
```{r}
  worstr <-
  specimens %>%
  add_count(class, binomial, name = "n") %>%
  add_count(class, binomial, sex, name = "nn") %>%
  select(class, binomial, sex, n, nn) %>%
  # Add in data for species with no males or no females
  complete(sex, nesting(class, binomial, n), fill = list(nn = 0)) %>%
  distinct() %>%
  mutate(percent = round(nn/n*100, 2)) %>%
  filter(sex == "Female" & n >= 10) %>%
  arrange(desc(class), percent)

  kable(slice(worstr, 1:10))
```

```{r}
  bestr <-
  specimens %>%
  add_count(class, binomial, name = "n") %>%
  add_count(class, binomial, sex, name = "nn") %>%
  select(class, binomial, sex, n, nn) %>%
  # Add in data for species with no males or no females
  complete(sex, nesting(class, binomial, n), fill = list(nn = 0)) %>%
  distinct() %>%
  mutate(percent = round(nn/n*100, 2)) %>%
  filter(sex == "Female" & n >= 10) %>%
  arrange(desc(class), -percent)

  kable(slice(bestr, 1:10))
```

## What about name bearing type specimens?
If we only consider name bearing type specimens the bias becomes more extreme. 

```{r}
  type_sum <-
  specimens %>%
  filter(type == "Type") %>%
  add_count(class, name = "n") %>%
  add_count(class, sex, name = "nn") %>%
  select(class, sex, n, nn) %>%
  distinct() %>%
  mutate('%' = round(nn/n*100, 2))

kable(type_sum)
```

### What if we split by the kind of type?

```{r}
  type_sum2 <-
  specimens %>%
  filter(type != "NonType") %>%
  add_count(class, type, name = "n") %>%
  add_count(class, sex, type, name = "nn") %>%
  select(class, sex, type, n, nn) %>%
  distinct() %>%
  mutate('%' = round(nn/n*100, 2))

kable(type_sum2)
```


```{r}
  type_sum3 <-
  specimens %>%
  filter(!is.na(typeStatus)) %>%
  add_count(class, typeStatus, name = "n") %>%
  add_count(class, sex, typeStatus, name = "nn") %>%
  select(class, sex, typeStatus, n, nn) %>%
  distinct() %>%
  mutate('%' = round(nn/n*100, 2))

kable(type_sum3)
```
