---
title: "Specimens per species - supplemental analysis"
output:
  html_document: default
---

```{r}
# Preliminaries

# Load libraries
library(tidyverse)
library(patchwork) # from GitHub
library(png)
library(knitr)
library(broom)
library(here)
library(ggfortify)

# Colour blind friendly palette
cbPalette <- c("#a53606", "#b32db5", "#881a58", "#0e288e", "#164c64")


```

```{r}
# Read in all the specimen data
specimens <- read_csv(here("data/all-specimen-data-2021-07.csv")) 
```

## How does the proportion of male and female specimens vary as we increase the number of specimens?
We expect large skews in the percentage of male or female specimens when sample size is low, but the ratio of male:female specimens should approach 50:50 as more specimens are added (assuming a 50:50 sex ratio in nature). 

Most species are repesented by only a few specimens (see histograms below), with large skews in % female specimens (in both directions) at low numbers.

```{r, fig.cap = "Distribution of percentage female specimens and log number of specimens per species across herps."}
source(here("analyses/figure-specimens-all.R"))
hist1 / hist2
```

```{r, fig.cap = "Relationship between the percentage female specimens per species and the number of specimens per species. Hex bins are used rather than points to make the plot clearer."}
all
```

If we fit a model (LM or GLM) to this data we see massive heteroscedasticity in the residuals.
```{r, fig.cap = "Model diagnostic plots for models of % female specimens as a function of log total number of specimens for each species"}
ds <-
  specimens %>%
  filter(!is.na(sex)) %>%
  group_by(class) %>%
  add_count(binomial, name = "n") %>%
  add_count(binomial, sex, name = "nn") %>%
  # Add in data for species with no males or no females
  complete(sex, nesting(class, binomial, n), fill = list(nn = 0)) %>%
  filter(sex == "Female") %>%
  rename(female = nn) %>%
  mutate(male = n - female) %>%
  select(class, order, binomial, n, female, male) %>%
  mutate(percentf = (female/n) * 100) %>%
  distinct()

spec_model <- glm(cbind(female, male) ~ log(n+1), data = ds, family = "binomial")

# Diagnostics
autoplot(spec_model, add.smooth = FALSE)
```

And we see a significant relationship between number of specimens and proportion female/male
```{r}
# Is the relationship significant?
anova(spec_model, test = "Chisq")
```

To avoid this variability influencing our results we excluded all species with fewer than 10 specimens from our analyses. Model fit improved a little (see below).

```{r, fig.cap = "Model diagnostic plots for models of % female specimens as a function of log total number of specimens for each species, for species with > 10 specimens"}
ds10 <- 
  ds %>%
  filter(n >= 10)

spec_model2 <- glm(cbind(female, male) ~ log(n+1), data = ds10, family = "binomial")

# Diagnostics
autoplot(spec_model2, add.smooth = FALSE)
```

And we no longer see a significant relationship between number of specimens and proportion female/male
```{r}
# Is the relationship significant?
anova(spec_model2, test = "Chisq")
```
